Compositional equivalences based on Open pNets

### 2020-09-03 2016-Weak-FH-Bisim/WeakBisim.tex
quelques states variables, et autres details...

A relation over the states of two open automata  $<\!J,\mathcal{S}_1, s_0,V_1,
   \mathcal{T}_1\!>$ and $<\!J,\mathcal{S}_2,t_0,V_2, \mathcal{T}_2\!>$ has the form $\mathcal{R}=\{(s,t|\Pred_{s,t})\}$, it relates states of $\mathcal{S}_1$ and 
$\mathcal{S}_2$ constrained by a predicate.
More precisely, for any pair $(s,t)\in \mathcal{S}_1\times \mathcal{S}_2$, there is a 
   single
      $(s,t|\Pred_{s,t})\in\mathcal{R}$  stating that $s$ and $t$ are related 
      if $\Pred_{s,t}$       is 
      True, i.e. the states are related when the variables in $V_1$ and $V_2$ verify the 
      predicate $\Pred_{s,t}$. By nature, this is well-defined if the variables of the two open automata are disjoint, i.e.  $V_1\cap V_2=\emptyset$.
 FH-bisimulation is defined formally: 
 \begin{definition}[Strong FH-bisimulation]\label{def-FH-bisim} ~\\
\noindent
\begin{minipage}{0.69\linewidth} 	Suppose 
   $A_1 = <\!J,\mathcal{S}_1, s_0,V_1,
   \mathcal{T}_1\!>$ and $A_2 = <\!J,\mathcal{S}_2,t_0,V_2, \mathcal{T}_2\!>$
   are open automata with identical holes of the same sort, with disjoint state variables.

@@

    Figure \ref{SimpleProtCounter:ImplOA} shows the open automaton of the implementation $SimpleSystem$ (Figure \ref{SimpleProt:Impl}). In this drawing, we have short labels for states, representing $\ostate{s_0,m_0,r_0}$ by \texttt{000}. Note that open transitions are denoted $\texttt{SI}_i$ and tau open transition by  $\texttt{SI}_{\tau}$. The resulting behaviour is quite simple:  we have a main loop including receiving a message from $P$ and transmitting the same message to $Q$, with some intermediate $\tau$ actions from the internal communications between the protocol processes. In most of the transitions, you can observe that data is propagated between the successive state variables (holding the message, and the error counter value). On the right of the figure, you have a loop of $\tau$ actions ($\texttt{SI}_4$, $\texttt{SI}_5$ and $\texttt{SI}_6$)  showing the handling of errors and corresponding increments of the error counter.

@@

 For readability, we do not include the full details of the Implementation weak open transitions in Figure \ref{SimpleProtCounter:ImplWOA2}, but we show some examples below and the full list is included in Appendix \ref{Appendix:FullExample}.
First, let us point out that the weak OT loops ($WI_1$,$WI_2$ and $W_\tau$) on state ${000}$ are also present in all other states, we did not repeat them. Then many WOTs are similar, and numbered accordingly as 3, 3a, 3b, 3c and 8, 8a, 8b, 8c respectively: they only differ by the name of the state variables in their respective source or target states; the "variant" WOTs appear in blue in   Figure \ref{SimpleProtCounter:ImplWOA2}.
\end{example}

@@

\begin{definition}[Weak FH-bisimulation]\label{def-Weak-bisim} ~\\
\noindent
Let $A_1 = <J,\mathcal{S}_1, s_0,V_1,
    \mathcal{T}_1>$ and $A_2 = <J,\mathcal{S}_2,t_0, V_2, \mathcal{T}_2>$ be open automata with disjoint state variables.
Let $<J,\mathcal{S}_1, s_0,V_1,
    \WT_1>$ and $<J,\mathcal{S}_2,t_0, V_2, \WT_2>$ be the
weak open automata derived from $A_1$ and $A_2$ respectively.
Let $\mathcal{R}$ a relation over
$\mathcal{S}_1$ and $\mathcal{S}_2$, as in Definition~\ref{def-FH-bisim}.

@@

This drawing is based on the observation that states 202 and 000 are only linked by a "pure $\tau$" transition, and have exactly the same possible behaviours.
More tricky, they do not have the same state variables, as 202 has the variables of state $s2$, but none of these variables are used in transitions from 202.
In this configuration we can guarantee that they are Weak bisimilar, and we have merged their (incoming and outgoing) transitions in the Figure. We denote this 
equivalence class of states as $\{000,202\}$.

>>>>

A relation over the states of two open automata  $<\!J,\mathcal{S}_1, s_0,V_1,
   \mathcal{T}_1\!>$ and $<\!J,\mathcal{S}_2,t_0,V_2, \mathcal{T}_2\!>$ has the form $\mathcal{R}=\{(s,t|\Pred_{s,t})\}$, it relates states of $\mathcal{S}_1$ and 
$\mathcal{S}_2$ constrained by a predicate.
More precisely, for any pair $(s,t)\in \mathcal{S}_1\times \mathcal{S}_2$, there is a 
   single
      $(s,t|\Pred_{s,t})\in\mathcal{R}$  stating that $s$ and $t$ are related 
      if $\Pred_{s,t}$       is 
      True, i.e. the states are related when the variables in $V_1$ and $V_2$ verify the 
      predicate $\Pred_{s,t}$. By nature, this is well-defined if the variables of the two open automata are disjoint, i.e.  $V_1\cap V_2=\emptyset$.
 FH-bisimulation is defined formally: 
 \begin{definition}[Strong FH-bisimulation]\label{def-FH-bisim} ~\\
\noindent
\begin{minipage}{0.69\linewidth} 	Suppose 
   $A_1 = <\!J,\mathcal{S}_1, s_0,V_1,
   \mathcal{T}_1\!>$ and $A_2 = <\!J,\mathcal{S}_2,t_0,V_2, \mathcal{T}_2\!>$
   are open automata with identical holes of the same sort, with disjoint sets of variables.

@@

    Figure \ref{SimpleProtCounter:ImplOA} shows the open automaton of the implementation $SimpleSystem$ (Figure \ref{SimpleProt:Impl}). In this drawing, we have short labels for states, representing $\ostate{s_0,m_0,r_0}$ by \texttt{000}. Note that open transitions are denoted $\texttt{SI}_i$ and tau open transition by  $\texttt{SI}_{\tau}$. The resulting behaviour is quite simple:  we have a main loop including receiving a message from $P$ and transmitting the same message to $Q$, with some intermediate $\tau$ actions from the internal communications between the protocol processes. In most of the transitions, you can observe that data is propagated between the successive plts variables (holding the message, and the error counter value). On the right of the figure, you have a loop of $\tau$ actions ($\texttt{SI}_4$, $\texttt{SI}_5$ and $\texttt{SI}_6$)  showing the handling of errors and corresponding increments of the error counter.

@@

 For readability, we do not include the full details of the Implementation weak open transitions in Figure \ref{SimpleProtCounter:ImplWOA2}, but we show some examples below and the full list is included in Appendix \ref{Appendix:FullExample}.
First, let us point out that the weak OT loops ($WI_1$,$WI_2$ and $W_\tau$) on state ${000}$ are also present in all other states, we did not repeat them. Then many WOTs are similar, and numbered accordingly as 3, 3a, 3b, 3c and 8, 8a, 8b, 8c respectively: they only differ by the name of the \TODO{ state variables} in their respective source or target states; the "variant" WOTs appear in blue in   Figure \ref{SimpleProtCounter:ImplWOA2}.
\end{example}

@@

\begin{definition}[Weak FH-bisimulation]\label{def-Weak-bisim} ~\\
\noindent
Let $A_1 = <J,\mathcal{S}_1, s_0,V_1,
    \mathcal{T}_1>$ and $A_2 = <J,\mathcal{S}_2,t_0, V_2, \mathcal{T}_2>$ be open automata with disjoint sets of variables.
Let $<J,\mathcal{S}_1, s_0,V_1,
    \WT_1>$ and $<J,\mathcal{S}_2,t_0, V_2, \WT_2>$ be the
weak open automata derived from $A_1$ and $A_2$ respectively.
Let $\mathcal{R}$ a relation over
$\mathcal{S}_1$ and $\mathcal{S}_2$, as in Definition~\ref{def-FH-bisim}.

@@

This drawing is based on the observation that states 202 and 000 are only linked by a "pure $\tau$" transition, and have exactly the same possible behaviours.
%More tricky, they do not have the same state variables, as 202 has the variables of state $s2$, but none of these variables are used in transitions from 202.
In this configuration we can guarantee that they are Weak bisimilar, and we have merged their (incoming and outgoing) transitions in the Figure. We denote this 
equivalence class of states as $\{000,202\}$.
